A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-08-15 tidyindex.qmd
rework distribution fit/ benchmark/ simplification

| No. | Module                  | Input                                                                  | Operation                                                                         | Output                                                                 |
|-|-------------------|--------------|----------------------|----------------------|
| 1   | Temporal processing     | $x^i_{s}(t)$                                                           | $f[x^i_s(t)]$                                                                         | $\tilde{x}^i_s(t^\prime)$, where $t^\prime \in D_{t^\prime}$            |
| 2   | Spatial processing      | $x^i_t(s)$                                                             | $g[x^i_t(s)]$                                                                         | $x^i_t(s^\prime)$, where $s^\prime \in D_{s^\prime}$ ????                   |
| 3   | Variable transformation | $x^i(s; t)$                                                            | $T[x^i(s;t)]$                                                                       | $Tx^i(s;t)$ ????                                                       |
| 4   | Scaling                 | $x^i(s; t)$                                                            | $[x^i(s;t) - \alpha]/\gamma$                                                        | $Sx^i(s;t)$ ????                                                       |
| 5   | Dimension reduction     | $\mathbf{x}(s;t)$, where $\mathbf{x} \subseteq \mathbb{R}^q,\; q\leq p$ | $h[\mathbf{x}(s;t)]$                                                              | $\mathbf{y}(s;t)$, where $\mathbf{y} \subseteq \mathbb{R}^d$ and $d < q$ |
| 6   | Distribution fit        | $x^i(s; t)$                                                            | $F[x^i(s;t)]$, where $F$ is the CDF function | $p^i(s;t)$, where $p(.) \in [0, 1]$ is a probability value                |
| 7   | Normalising             | $x^i(s; t)$                                                            | $\Phi^{-1}[x^i(s; t)]$                                                              | $z^i(s; t)$, where $z$ has a normal distribution                          |
| 8   | Benchmarking            | $x^i(s; t)$                                                            | $u[x^i(s;t)]$                                                                       | $b^i(s;t)$                                                               |
| 9   | Simplification          | $x^i(s; t)$                                                            | $v[x^i(s;t)]$                                                                            | $A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}$                                 |

@@

<!-- Other weight constraint used includes $\sum_{p=1}^P\lambda_p^2 = 1$, is used by principal component analysis.  -->

<!-- Spectral satellite imageries often collected at different bands and it is common to combine values from multiple bands to compose new variables relevant to the index. The most common example of this is the Normalized Difference Vegetation Index (NDVI), which uses near-infrared (NIR) ($x_1(s;t)$) and red channel ($x_2(s;t)$): -->

<!-- \begin{equation} -->

<!-- \text{NDVI}(s;t) = \frac{x_1(s;t) - x_2(s;t)}{x_1(s;t) + x_2(s;t)}, -->

<!-- \end{equation} -->

<!-- which is used to evaluate the vegetation from near-infrared (NIR) and red channel. -->

@@

<!-- *model fit? * -->

`Input`: $x^l(s; t)$

`Output`:

`Computation`: Distribution fit can be seen as the model fitting in its simplest term. It can be represented by

```{=tex}
\begin{equation}
F_{\eta}(x(\mathbf{s}; \mathbf{t})), 
\end{equation}
```
where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. A distribution fit typically aims at finding the distribution that best fits the data. Analysts may start from a pool of candidate distributions with a chosen fitting method and goodness of fit measure. While it is useful to find the ultimate best distribution to fits the data, from a probabilistic perspective, the fitting procedure itself has an uncertainty associated with the data fed and the parameter chosen. A reasonable alternative is to understand how much the index values can vary given different distributions, fitting methods, and goodness of fit tests, and whether these variations are negligible in a given application.

@@

`Input`: $x^l(s; t)$

`Output`:

`Computation`: This step maps the univariate series into a different scale, typically for ease of comparison across regions. For example, a normal scale, \[0, 1\], or \[0, 100\] may be favored for reporting certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution are converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$. Normalising is usually used at the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of a normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence. This is concerning, especially at the extreme values, since a small difference in the tail density can have magnitudes of difference in its probability of occurrence.

```{r fig-normalising}
#| fig-cap: "Scatterplot of normal quantiles against their density values. THree tail density values are highlighted with its probability of occurence labelled. Probability is calculated assuming monthly data: with a density of -2, the probability of occurrence is 1/pnorm(-2)/12 = 4 years. The non-linear relationship between the two quantities suggests normalised indexes need to be interpreted with caution since a slight change in the tail distribution can result in magnitudes of difference in its probability of occurrence."
#| fig-width: 9
library(tidyverse)
all <- tibble(x = seq(-3, 3, 0.01), y = pnorm(x)) 
dt <- tibble(x = c(-3, -2.5, -2), y = pnorm(x)) 
label <- tibble(x = c(-3, -2.5, -2), y = pnorm(x), 
                label = c("once in 62 years", "once in 13 years", "once in 4 years"))

dt %>% 
  ggplot() + 
  geom_point(aes(x= x, y = y)) + 
  geom_line(data = all, aes(x = x, y = y)) + 
  ggrepel::geom_label_repel(
    data = label, aes(x = x, y = y, label =label),
    nudge_y = 0.2, nudge_x = 0.1,
    arrow = arrow(length = unit(0.08, "inches")), min.segment.length = 0) +
  scale_x_continuous(breaks = seq(-3, 3, 1)) + 
  xlab("Normalised Score") + 
  ylab("Probability") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank())
```

@@

`Input`: $x^l(s; t)$

`Output`:

$D(x^l(s,t), a)$ where $D(., .)$ is a distance function between

`Computation`: Benchmarking sets a constant value to allow the constructed index to be compared across time. Here we denote it with $u[x(s_i, t_j)]$ where $u$ is a scalar of interest in the index constructed. A benchmark value could be a constant or a function of the data, i.e. mean.

@@

`Input`: $x^l(s; t)$

`Output`: $C^l(s;t)$ where $\{C: c_0, c_1, \cdots, c_z\}$ is a discrete set

`Computation`: In public communication, the index values are usually accompanied by a categorical grade. The categorised grades are an ordered set of descriptive words or colors to communicate the severity or guide the comprehension of the indexes. The mapping from continuous index values to the discrete grades is called simplification in the pipeline and it can be written as a piece-wise function:

```{=tex}
\begin{equation}
v[x_i(s;t)] = 
\begin{cases}
a_0 & C_1 \leq x(s; t) < C_0 \\
a_1 & C_2 \leq x(s; t) < C_1 \\
a_2 & C_3 \leq x(s; t) < C_2 \\
\cdots \\
a_z & C_z \leq x(s; t)
\end{cases}
\end{equation}
```
where $C_0, C_1,\cdots ,C_z$ are the categories and $c_0, c_1, \cdots, c_z$ are the thresholds for each category. In SPI, droughts are sorted into four categories: mild drought: $[-0.99, 0]$; moderate drought: $[-1.49, -1]$; severe drought: $[-1.99, -1.5]$, and extreme drought: $[-\infty, -2]$. In this case, $C_0, C_1, C_2, C_3$ are the drought categories: mild, moderate, severe, and extreme drought ($z = 3$) and $c_0 =0, c_1 = -1, c_2 = -1.5, c_3 = -2$ are the cutoff value for each class.

>>>>

| No. | Module                  | Input                                                                  | Operation                                                                         | Output                                                                 |
|-|-------------------|--------------|----------------------|----------------------|
| 1   | Temporal processing     | $x^i_{s}(t)$                                                           | $f[x^i_s(t)]$                                                                         | $\tilde{x}^i_s(t^\prime)$, where $t^\prime \in D_{t^\prime}$            |
| 2   | Spatial processing      | $x^i_t(s)$                                                             | $g[x^i_t(s)]$                                                                         | $x^i_t(s^\prime)$, where $s^\prime \in D_{s^\prime}$ ????                   |
| 3   | Variable transformation | $x^i(s; t)$                                                            | $T[x^i(s;t)]$                                                                       | $Tx^i(s;t)$ ????                                                       |
| 4   | Scaling                 | $x^i(s; t)$                                                            | $[x^i(s;t) - \alpha]/\gamma$                                                        | $Sx^i(s;t)$ ????                                                       |
| 5   | Dimension reduction     | $\mathbf{x}(s;t)$, where $\mathbf{x} \subseteq \mathbb{R}^q,\; q\leq p$ | $h[\mathbf{x}(s;t)]$                                                              | $\mathbf{y}(s;t)$, where $\mathbf{y} \subseteq \mathbb{R}^d$ and $d < q$ |
| 6   | Distribution fit        | $x^i(s; t)$                                                            | $F[x^i(s;t)]$, where $F$ is the CDF | $p^i(s;t)$, where $p(.) \in [0, 1]$ is a probability value                |
| 7   | Normalising             | $x^i(s; t)$                                                            | $\Phi^{-1}[x^i(s; t)]$                                                              | $z^i(s; t)$, where $z$ has a normal distribution                          |
| 8   | Benchmarking            | $x^i(s; t)$                                                            | $u[x^i(s;t)]$                                                                       | $b^i(s;t)$                                                               |
| 9   | Simplification          | $x^i(s; t)$                                                            | $v[x^i(s;t)]$                                                                            | $A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}$                                 |

@@

@@

Distribution fit applies the cumulative distribution function $F$ of a distribution on the variable $x^i(s; t)$ to obtain the probability values $p^i(s;t) \in [0, 1]$. In SPEI, many distributions, including log-logistic, pearson III, lognormal, and general extreme distribution, are candidates for the aggregated series. Different fitting methods, and different goodness of fit tests may be used to compare the distribution choice on the index value.

@@

<!-- This step maps the univariate series into a different scale, typically for ease of comparison across regions. For example, a normal scale, \[0, 1\], or \[0, 100\] may be favored for reporting certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution are converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$. Normalising is usually used at the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of a normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence. This is concerning, especially at the extreme values, since a small difference in the tail density can have magnitudes of difference in its probability of occurrence. -->

@@

Benchmark sets a value $b^i(s,t)$ for comparing against the original variable $x^i(s;t)$. This benchmark can be a fixed value consistently across space and time or determined by the data through the function $u[x^i(s;t)]$. Once a benchmark being set, observations can be highlighted for adjustments in other modules or can serve targets for monitoring and planning.

@@

Simplification takes a continuous variable $x^i(s;t)$ and categorises it into a discrete set $A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}$ through a piecewise function,

```{=tex}
\begin{equation}
v[x_i(s;t)] = 
\begin{cases}
a_0 & C_1 \leq x^i(s; t) < C_0 \\
a_1 & C_2 \leq x^i(s; t) < C_1 \\
a_2 & C_3 \leq x^i(s; t) < C_2 \\
\cdots \\
a_z & C_z \leq x^i(s; t)
\end{cases}
\end{equation}
```

This is typically used at the end of the index pipeline to simplify the index to communicate to the public the severity of the concept of interest measured by the index. An example of simplification is to map the calculated SPI to four categories: mild, moderate, severe, and extreme drought.

### 2023-08-15 tidyindex.tex
rework distribution fit/ benchmark/ simplification

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

@@

\hypertarget{tbl-notation}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0128}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2436}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1795}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2821}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2821}}@{}}
\caption{\label{tbl-notation}An notation overview of the input,
operation, and output of each pipeline module.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endhead
1 & Temporal processing & \(x^i_{s}(t)\) & \(f[x^i_s(t)]\) &
\(\tilde{x}^i_s(t^\prime)\), where \(t^\prime \in D_{t^\prime}\) \\
2 & Spatial processing & \(x^i_t(s)\) & \(g[x^i_t(s)]\) &
\(x^i_t(s^\prime)\), where \(s^\prime \in D_{s^\prime}\) ???? \\
3 & Variable transformation & \(x^i(s; t)\) & \(T[x^i(s;t)]\) &
\(Tx^i(s;t)\) ???? \\
4 & Scaling & \(x^i(s; t)\) & \([x^i(s;t) - \alpha]/\gamma\) &
\(Sx^i(s;t)\) ???? \\
5 & Dimension reduction & \(\mathbf{x}(s;t)\), where
\(\mathbf{x} \subseteq \mathbb{R}^q,\; q\leq p\) &
\(h[\mathbf{x}(s;t)]\) & \(\mathbf{y}(s;t)\), where
\(\mathbf{y} \subseteq \mathbb{R}^d\) and \(d < q\) \\
6 & Distribution fit & \(x^i(s; t)\) & \(F[x^i(s;t)]\), where \(F\) is
the CDF function & \(p^i(s;t)\), where \(p(.) \in [0, 1]\) is a
probability value \\
7 & Normalising & \(x^i(s; t)\) & \(\Phi^{-1}[x^i(s; t)]\) &
\(z^i(s; t)\), where \(z\) has a normal distribution \\
8 & Benchmarking & \(x^i(s; t)\) & \(u[x^i(s;t)]\) & \(b^i(s;t)\) \\
9 & Simplification & \(x^i(s; t)\) & \(v[x^i(s;t)]\) &
\(A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}\) \\
\bottomrule()
\end{longtable}

@@

\texttt{Input}: \(x^l(s; t)\)

\texttt{Output}:

\texttt{Computation}: Distribution fit can be seen as the model fitting
in its simplest term. It can be represented by

\begin{equation}
F_{\eta}(x(\mathbf{s}; \mathbf{t})), 
\end{equation}

where \(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\). A
distribution fit typically aims at finding the distribution that best
fits the data. Analysts may start from a pool of candidate distributions
with a chosen fitting method and goodness of fit measure. While it is
useful to find the ultimate best distribution to fits the data, from a
probabilistic perspective, the fitting procedure itself has an
uncertainty associated with the data fed and the parameter chosen. A
reasonable alternative is to understand how much the index values can
vary given different distributions, fitting methods, and goodness of fit
tests, and whether these variations are negligible in a given
application.

@@

\texttt{Input}: \(x^l(s; t)\)

\texttt{Output}:

\texttt{Computation}: This step maps the univariate series into a
different scale, typically for ease of comparison across regions. For
example, a normal scale, {[}0, 1{]}, or {[}0, 100{]} may be favored for
reporting certain indexes. In drought indexes, i.e.~SPI or SPEI, the
quantiles from the fitted distribution are converted into the normal
scale via the normal reverse CDF function: \(\Phi^{-1}(.)\). Normalising
is usually used at the end of the pipeline and its main difference from
the scaling step is that here the change of scale also changes the
distribution of the variable. While being commonly used, this step can
get criticism from analysts for forcing the data into the decided scale,
which can be either unnecessary or inaccurately exaggerate or downplay
the outliers. Also, the use of a normal scale needs to be interpreted
with caution. Figure~\ref{fig-normalising} illustrates the normal
density not being directly proportional to its probability of
occurrence. This is concerning, especially at the extreme values, since
a small difference in the tail density can have magnitudes of difference
in its probability of occurrence.

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-normalising-1.pdf}

\caption{\label{fig-normalising}Scatterplot of normal quantiles against
their density values. THree tail density values are highlighted with its
probability of occurence labelled. Probability is calculated assuming
monthly data: with a density of -2, the probability of occurrence is
1/pnorm(-2)/12 = 4 years. The non-linear relationship between the two
quantities suggests normalised indexes need to be interpreted with
caution since a slight change in the tail distribution can result in
magnitudes of difference in its probability of occurrence.}

@@

\texttt{Input}: \(x^l(s; t)\)

\texttt{Output}:

\(D(x^l(s,t), a)\) where \(D(., .)\) is a distance function between

\texttt{Computation}: Benchmarking sets a constant value to allow the
constructed index to be compared across time. Here we denote it with
\(u[x(s_i, t_j)]\) where \(u\) is a scalar of interest in the index
constructed. A benchmark value could be a constant or a function of the
data, i.e.~mean.

@@

\texttt{Input}: \(x^l(s; t)\)

\texttt{Output}: \(C^l(s;t)\) where \(\{C: c_0, c_1, \cdots, c_z\}\) is
a discrete set

\texttt{Computation}: In public communication, the index values are
usually accompanied by a categorical grade. The categorised grades are
an ordered set of descriptive words or colors to communicate the
severity or guide the comprehension of the indexes. The mapping from
continuous index values to the discrete grades is called simplification
in the pipeline and it can be written as a piece-wise function:

\begin{equation}
v[x_i(s;t)] = 
\begin{cases}
a_0 & C_1 \leq x(s; t) < C_0 \\
a_1 & C_2 \leq x(s; t) < C_1 \\
a_2 & C_3 \leq x(s; t) < C_2 \\
\cdots \\
a_z & C_z \leq x(s; t)
\end{cases}
\end{equation}

where \(C_0, C_1,\cdots ,C_z\) are the categories and
\(c_0, c_1, \cdots, c_z\) are the thresholds for each category. In SPI,
droughts are sorted into four categories: mild drought: \([-0.99, 0]\);
moderate drought: \([-1.49, -1]\); severe drought: \([-1.99, -1.5]\),
and extreme drought: \([-\infty, -2]\). In this case,
\(C_0, C_1, C_2, C_3\) are the drought categories: mild, moderate,
severe, and extreme drought (\(z = 3\)) and
\(c_0 =0, c_1 = -1, c_2 = -1.5, c_3 = -2\) are the cutoff value for each
class.

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{tbl-notation}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0128}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2436}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1795}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2821}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2821}}@{}}
\caption{\label{tbl-notation}An notation overview of the input,
operation, and output of each pipeline module.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
No.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Input
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Output
\end{minipage} \\
\midrule()
\endhead
1 & Temporal processing & \(x^i_{s}(t)\) & \(f[x^i_s(t)]\) &
\(\tilde{x}^i_s(t^\prime)\), where \(t^\prime \in D_{t^\prime}\) \\
2 & Spatial processing & \(x^i_t(s)\) & \(g[x^i_t(s)]\) &
\(x^i_t(s^\prime)\), where \(s^\prime \in D_{s^\prime}\) ???? \\
3 & Variable transformation & \(x^i(s; t)\) & \(T[x^i(s;t)]\) &
\(Tx^i(s;t)\) ???? \\
4 & Scaling & \(x^i(s; t)\) & \([x^i(s;t) - \alpha]/\gamma\) &
\(Sx^i(s;t)\) ???? \\
5 & Dimension reduction & \(\mathbf{x}(s;t)\), where
\(\mathbf{x} \subseteq \mathbb{R}^q,\; q\leq p\) &
\(h[\mathbf{x}(s;t)]\) & \(\mathbf{y}(s;t)\), where
\(\mathbf{y} \subseteq \mathbb{R}^d\) and \(d < q\) \\
6 & Distribution fit & \(x^i(s; t)\) & \(F[x^i(s;t)]\), where \(F\) is
the CDF & \(p^i(s;t)\), where \(p(.) \in [0, 1]\) is a probability
value \\
7 & Normalising & \(x^i(s; t)\) & \(\Phi^{-1}[x^i(s; t)]\) &
\(z^i(s; t)\), where \(z\) has a normal distribution \\
8 & Benchmarking & \(x^i(s; t)\) & \(u[x^i(s;t)]\) & \(b^i(s;t)\) \\
9 & Simplification & \(x^i(s; t)\) & \(v[x^i(s;t)]\) &
\(A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}\) \\
\bottomrule()
\end{longtable}

@@

Distribution fit applies the cumulative distribution function \(F\) of a
distribution on the variable \(x^i(s; t)\) to obtain the probability
values \(p^i(s;t) \in [0, 1]\). In SPEI, many distributions, including
log-logistic, pearson III, lognormal, and general extreme distribution,
are candidates for the aggregated series. Different fitting methods, and
different goodness of fit tests may be used to compare the distribution
choice on the index value.

@@

@@

Benchmark sets a value \(b^i(s,t)\) for comparing against the original
variable \(x^i(s;t)\). This benchmark can be a fixed value consistently
across space and time or determined by the data through the function
\(u[x^i(s;t)]\). Once a benchmark being set, observations can be
highlighted for adjustments in other modules or can serve targets for
monitoring and planning.

@@

Simplification takes a continuous variable \(x^i(s;t)\) and categorises
it into a discrete set \(A^i(s;t) \in \{a_1, a_2, \cdots, a_j\}\)
through a piecewise function,

\begin{equation}
v[x_i(s;t)] = 
\begin{cases}
a_0 & C_1 \leq x^i(s; t) < C_0 \\
a_1 & C_2 \leq x^i(s; t) < C_1 \\
a_2 & C_3 \leq x^i(s; t) < C_2 \\
\cdots \\
a_z & C_z \leq x^i(s; t)
\end{cases}
\end{equation}

This is typically used at the end of the index pipeline to simplify the
index to communicate to the public the severity of the concept of
interest measured by the index. An example of simplification is to map
the calculated SPI to four categories: mild, moderate, severe, and
extreme drought.
